---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 7,
  fig.height = 5,
  dpi = 150
)
```

# bootmakr

<!-- badges: start -->
[![R-CMD-check](https://img.shields.io/badge/R--CMD--check-passing-brightgreen)](https://github.com/)
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

**Bootstrap inference for sensitivity analysis under omitted variable bias.**

`bootmakr` wraps the
[sensemakr](https://github.com/carloscinelli/sensemakr) package in a
bootstrap loop, producing bootstrap standard errors, percentile confidence
intervals, and *p*-values for the bias-adjusted treatment effect. It is the
companion R package to the Stata command of the same name.

## Why bootstrap the sensitivity bounds?

`sensemakr` computes analytical adjusted estimates and confidence intervals
under a hypothetical confounder with a given strength. These analytical
intervals rely on asymptotic OLS standard errors, which may not perform well
with clustered data, small samples, or complex survey designs. `bootmakr`
replaces the analytical inference with a nonparametric bootstrap — including
cluster and stratified bootstrap — so the CIs and *p*-values are robust to
these complications.

## Installation

```r
# Install from GitHub (once published):
# devtools::install_github("username/bootmakr")
```

## Quick start

```{r quickstart, message = FALSE}
library(bootmakr)

data(darfur, package = "sensemakr")

out <- bootmakr(
  peacefactor ~ directlyharmed + age + farmer_dar + herder_dar + pastvoted,
  data    = darfur,
  treat   = "directlyharmed",
  benchmark_covariates = "age",
  kd      = 1,
  reps    = 1000,
  seed    = 42,
  progress = FALSE
)
out
```

## Sweeping across benchmark strengths

Supply a vector of `kd` values to see how the adjusted effect changes as the
hypothetical confounder grows stronger. The `plot()` method produces a
coefficient plot with bootstrap CIs:

```{r kd-sweep, fig.height = 4.5}
out_sweep <- bootmakr(
  peacefactor ~ directlyharmed + age + farmer_dar + herder_dar + pastvoted,
  data    = darfur,
  treat   = "directlyharmed",
  benchmark_covariates = "age",
  kd      = seq(0.5, 3, by = 0.5),
  reps    = 1000,
  seed    = 42,
  progress = FALSE
)
out_sweep
plot(out_sweep, type = "kd_sweep")
```

## Cluster bootstrap

Pass a cluster identifier to `cluster` for a cluster-robust bootstrap — the
resampling is done at the cluster level:

```{r cluster}
# Simulated clustered data
set.seed(999)
G  <- rep(1:50, each = 20)
u  <- rnorm(50)[G]
x  <- 0.5 * u + rnorm(1000)
cv <- 0.3 * u + rnorm(1000)
y  <- 0.5 * x + 0.4 * cv + 0.8 * u + rnorm(1000)
sim <- data.frame(y = y, x = x, cv = cv, cid = G)

out_cl <- bootmakr(
  y ~ x + cv, data = sim, treat = "x",
  benchmark_covariates = "cv", kd = c(0.5, 1),
  reps = 1000, seed = 77, cluster = "cid",
  progress = FALSE
)
out_cl
```

## Grouped benchmarks

When the benchmark for the hypothetical confounder should reflect the *joint*
explanatory power of several covariates, use `gbenchmark_covariates`.
Internally this computes the group partial R² (via
`sensemakr::group_partial_r2`) and applies the proper nonlinear kd-scaling:

```{r grouped, eval = FALSE}
out_g <- bootmakr(
  lg_ceopay ~ owner_ceo + ceo_tenure + PA_nic3_med + ceo_edu_dummy +
    lg_sales + firm_age + promoters_pct + institutions_pct + year + nic,
  data    = ceo_data,
  treat   = "owner_ceo",
  gbenchmark_covariates = c("ceo_tenure", "ceo_edu_dummy"),
  kd      = seq(0.1, 0.5, by = 0.1),
  reps    = 10000,
  cluster = "firm_id",
  seed    = 912323
)
out_g
```

## Convergence diagnostics

Large-sample bootstrap inference depends on using enough replications. Pass
`converge = TRUE` (or a list with fine-grained control) to assess whether SEs
and *p*-values have stabilised:

```{r convergence, fig.height = 7}
out_conv <- bootmakr(
  peacefactor ~ directlyharmed + age + farmer_dar + herder_dar + pastvoted,
  data    = darfur,
  treat   = "directlyharmed",
  benchmark_covariates = "age",
  kd      = 1,
  reps    = 2000,
  seed    = 42,
  converge = list(minreps = 200, stepsize = 200, threshold = 1500),
  progress = FALSE
)
out_conv
plot(out_conv, type = "convergence")
```

## Accessing the raw bootstrap draws

All bootstrap replicates are stored in the returned object, so there is no
need to re-run the analysis to inspect the distribution:

```{r draws}
draws <- out$boot_samples[, 1]          # vector of B adjusted estimates
draws <- draws[is.finite(draws)]

# Percentiles, moments, etc.
quantile(draws, c(0.025, 0.5, 0.975))

# Or export for further analysis
# write.csv(data.frame(adjusted_estimate = draws), "boot_draws.csv")
```

## Key arguments

| Argument | Description |
|---|---|
| `formula`, `data`, `treat` | Standard OLS specification and treatment name |
| `benchmark_covariates` | Individual benchmark covariate(s) |
| `gbenchmark_covariates` | Grouped benchmark covariates (joint partial R²) |
| `kd`, `ky` | Benchmark strength multipliers (`ky` defaults to `kd`) |
| `reps`, `seed` | Number of bootstrap replications and random seed |
| `cluster`, `strata` | Cluster and/or strata identifiers |
| `alpha` | Significance level (default 0.05) |
| `converge` | `TRUE`, `FALSE`, or `list(minreps, stepsize, threshold)` |
| `progress` | Show a progress bar (default `TRUE`) |

## Methods

| Method | Description |
|---|---|
| `print(x)` | Coefficient table with bootstrap SEs, CIs, and *p*-values |
| `plot(x, type = "kd_sweep")` | Coefficient plot across kd values |
| `plot(x, type = "histogram")` | Bootstrap distribution histogram |
| `plot(x, type = "convergence")` | Three-panel convergence diagnostic plot |
| `plot(x)` | Auto-selects the most informative plot |

## References
Cinelli, C. and Hazlett, C. (2020). Making Sense of Sensitivity: Extending
Omitted Variable Bias. *Journal of the Royal Statistical Society, Series B
(Statistical Methodology)*, 82(1), 39–67.

Cinelli, C., Ferwerda, J. and Hazlett, C. (2024). sensemakr: Sensitivity analysis tools for OLS in R and Stata. *Observational Studies*, 10(2), 93-127.

Lonati, S. and Wulff, J. N. (2026). Why you should not use the ITCV with robust standard errors (and what to do instead). *SSRN Working Paper*.

## License

MIT
