---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 150
)
```

# bootmakr

<!-- badges: start -->
[![R-CMD-check](https://github.com/jespernwulff/bootmakr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jespernwulff/bootmakr/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

**Bootstrap inference for sensitivity analysis to omitted variables.**

**Documentation, a getting-started guide, a worked example and the argument
behind the package, in R and Stata:
<https://jespernwulff.github.io/bootmakr/>**

`bootmakr` wraps [sensemakr](https://carloscinelli.com/sensemakr/) (Cinelli
& Hazlett, 2020) in a bootstrap loop. For an omitted variable as strong as an
observed benchmark covariate, it computes the bias-adjusted coefficient and
obtains its standard error, percentile confidence interval and *p*-value by
resampling observations, whole clusters or units within strata. It is the
companion R package to Lonati and Wulff (2026) and to the
[Stata command of the same name](https://github.com/jespernwulff/bootmakr-stata).

## Why bootstrap?

The impact threshold of a confounding variable (ITCV) is read as the product
of correlations an omitted variable needs to overturn a result. Lonati and
Wulff (2026) show that this reading breaks down when the regression is
estimated with heteroskedasticity- or cluster-robust standard errors, and
that the analytic confidence interval of `sensemakr` breaks down for the
same reason. The remedy is the bootstrap that Cinelli, Ferwerda and Hazlett
(2024, Appendix C) propose: resample the data the way you would choose
standard errors, refit the regression and recompute the adjusted estimate in
every replication. `bootmakr` automates it, and reports how strong the
assumed omitted variable is on the scale of the ITCV. See
[Why bootstrap?](https://jespernwulff.github.io/bootmakr/articles/why-bootstrap.html).

## Installation

```r
# install.packages("remotes")
remotes::install_github("jespernwulff/bootmakr")
```

## Quick start

The package comes with `firms`, a simulated panel of 250 firms observed for
20 years. The true effect of `x` on `y` is 0.25. An unobserved firm-level
variable `q`, as strong as the observed control `c`, biases the regression
that omits it; the within-firm components of `x` and of the disturbance are
persistent, so standard errors must be clustered by firm. Here is what the
coefficient of `x` would be had a variable as strong as `c` been omitted,
with firms resampled as whole clusters:

```{r quick}
library(bootmakr)
data("firms")

out <- bootmakr(y ~ x + c, data = firms, treat = "x",
                benchmark_covariates = "c", cluster = "firm",
                reps = 1000, seed = 123, progress = FALSE)
out
```

The adjusted estimate is close to the true 0.25, and its bootstrap interval
excludes zero. The block below the table says how strong "as strong as `c`"
is: as partial R² values and as the *impact*, the product of the omitted
variable's partial correlations with the outcome and with the treatment,
which is the scale of the ITCV.

Several strengths at once show where the result breaks down:

```{r sweep}
sweep <- bootmakr(y ~ x + c, data = firms, treat = "x",
                  benchmark_covariates = "c", kd = seq(0.5, 1.5, by = 0.25),
                  cluster = "firm", reps = 1000, seed = 123, progress = FALSE)
sweep$results[, c("kd", "estimate", "se", "ci_lower", "ci_upper", "pvalue", "impact")]
plot(sweep, type = "kd_sweep")
```

Because the data are simulated, every answer can be checked against the
regression that includes `q`:

```{r check}
coef(summary(lm(y ~ x + c + q, data = firms)))["x", 1:2]
```

## Key arguments

| Argument | Description |
|---|---|
| `formula`, `data`, `treat` | Standard OLS specification and treatment name |
| `benchmark_covariates` | Individual benchmark covariate(s) |
| `gbenchmark_covariates` | Grouped benchmark covariates (joint partial R²) |
| `kd`, `ky` | Benchmark strength multipliers (`ky` defaults to `kd`) |
| `reps`, `seed` | Number of bootstrap replications and random seed |
| `cluster` | Cluster identifier(s). Length 1 → one-way cluster bootstrap with percentile CIs. Length 2 (e.g. `c("firm", "year")`) → two-way CGM bootstrap with normal-approx CIs |
| `strata` | Stratification identifier (not combined with two-way clustering) |
| `weights` | Regression weights (column name or vector) |
| `alpha` | Significance level (default 0.05) |
| `converge` | `TRUE`, `FALSE`, or `list(minreps, stepsize, threshold)` |
| `progress` | Show a progress bar (default `TRUE`) |

## Methods

| Method | Description |
|---|---|
| `print(x)` | Coefficient table with bootstrap SEs, CIs and *p*-values, and the benchmark-strength block |
| `plot(x, type = "kd_sweep")` | Coefficient plot across kd values |
| `plot(x, type = "histogram")` | Bootstrap distribution histogram |
| `plot(x, type = "convergence")` | Three-panel convergence diagnostic plot |
| `plot(x)` | Auto-selects the most informative plot |

## Citation

If you use `bootmakr`, please cite the paper it accompanies and the method
it builds on; `citation("bootmakr")` prints the entries.

## References

Cinelli, C. and Hazlett, C. (2020). Making Sense of Sensitivity: Extending
Omitted Variable Bias. *Journal of the Royal Statistical Society, Series B
(Statistical Methodology)*, 82(1), 39–67.
<https://doi.org/10.1111/rssb.12348>

Cinelli, C., J. Ferwerda, and C. Hazlett (2024). sensemakr: Sensitivity
analysis tools for OLS in R and Stata. *Observational Studies*, 10(2),
93–127. <https://doi.org/10.1353/obs.2024.a946583>

Lonati, S. and J. N. Wulff (2026). Why you should not use the ITCV with
robust standard errors (and what to do instead). *Academy of Management
Proceedings*, 2026(1). <https://doi.org/10.5465/AMPROC.2026.247bp>

## License

MIT
